A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-01-20 paper/paper.tex
paper: minor fixes after LiveShare session with Hugo

\begin{align*}
  \Dom & : \Type_m                                        \\
  \UIP & : \Pi x y : \Dom, \Pi p q : x = y, p = q
\end{align*}

@@

\begin{enumerate}
  \item[(i)] UIP holds on the unit and bool type:
    \begin{align*}
      \textsf{eq\_refl}        & : \Pi x, x = x                                                    \\
      \textsf{UIP\_refl\_unit} & : \Pi u : \textsf{unit}, \Pi x : u = u, x = \textsf{eq\_refl} \\
      \textsf{UIP\_refl\_bool} & : \Pi b : \textsf{bool}, \Pi x : b = b, x = \textsf{eq\_refl}
    \end{align*}
  \item[(ii)] UIP propagates to $\Sigma$-types.
  \item[(iii)] UIP propagates to $\Pi$-types, with some additional functional extensionality axioms.
\end{enumerate}

@@

\begin{align*}
  \textsf{UIP\_refl} & : \Pi x:\Dom, \Pi p : x = x, p = \textsf{eq\_refl } x
\end{align*}

@@

\begin{align*}
  X_0 &\defeq X:& \U \\
  X_1 &\defeq X_\kstar:& {\U}_\kstar(X)(X) \\
  X_2 &\defeq X_{\kstar\kstar}:& ({\U}_\kstar(X)(X))_{\kstar}(X_\kstar)(X_\kstar)
\end{align*}
of parametricity applied to an initial $X: \U$. This tells us how the previous (informal) type given to $X_2$ could be rephrased so that it comes as the instance of a general recipe characterizing the type of all $X_i$.

@@

\section{Perspectives}

@@

By equipping the universe construction with a structure of equivalences (as suggested along the lines of Altenkirch and Kaposi~\cite{}), as well as connections~\cite{}, we expect the parametric model to extend into a model of cubical type theory~\cite{}. A key point in this approach, is that, thanks to the indexed construction, we expect univalence to hold by definition in the model, without requiring any kind of Reedy fibrancy as it would a priori be needed in the fibered case. If these intuitions are correct, this would definitively root cubical type theory on top of iterated parametricity translation.

>>>>

\begin{align*}
  \Dom & : \Type_m                                  \\
  \UIP & : \Pi x y: \Dom.\, \Pi p q: x = y.\, p = q
\end{align*}

@@

\begin{enumerate}
  \item[(i)] UIP holds on the unit and bool type:
    \begin{align*}
      \textsf{eq\_refl}        & : \Pi x.\, x = x                                             \\
      \textsf{UIP\_refl\_unit} & : \Pi u: \unittype.\, \Pi x : u = u.\, x = \textsf{eq\_refl} \\
      \textsf{UIP\_refl\_bool} & : \Pi b: \booltype.\, \Pi x : b = b.\, x = \textsf{eq\_refl}
    \end{align*}
  \item[(ii)] UIP propagates to $\Sigma$-types.
  \item[(iii)] UIP propagates to $\Pi$-types, with some additional functional extensionality axioms.
\end{enumerate}

@@

\begin{align*}
  \textsf{UIP\_refl} & : \Pi x: \Dom.\, \Pi p: x = x.\, p = \textsf{eq\_refl } x
\end{align*}

@@

\begin{align*}
  X_0 & \defeq X                & :\, & \U                                               \\
  X_1 & \defeq X_\kstar         & :\, & {\U}_\kstar(X)(X)                                \\
  X_2 & \defeq X_{\kstar\kstar} & :\, & ({\U}_\kstar(X)(X))_{\kstar}(X_\kstar)(X_\kstar)
\end{align*}
of parametricity applied to an initial $X: \U$. This tells us how the previous (informal) type given to $X_2$ could be rephrased so that it comes as the instance of a general recipe characterizing the type of all $X_i$.

@@

\section{Future work}

@@

By equipping the universe construction with a structure of equivalences , as suggested along the lines of Altenkirch and Kaposi~\cite{}, as well as connections~\cite{}, we expect the parametric model to extend into a model of cubical type theory~\cite{}. A key point in this approach, is that, thanks to the indexed construction, we expect univalence to hold by definition in the model, without requiring any kind of Reedy fibrancy as it would a priori be needed in the fibered case. If these intuitions are correct, this would definitively root cubical type theory on top of iterated parametricity translation.
